---
title: "File Management Basics - Dropbox_Example"
author: "Tom Tibbett"
output:
  html_document:
    keep_md: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

### Unit 3: Calling Data from a Dropbox Account

Objectives

* Downloading Data from Christopher Gandrud's Dropbox Folder
* Using rio package to import something from URLs
* File is on Financial Regulators

Installing rio package.  IMPORTANT: NEED install_formats() to complete the installation

* install.packages("rio")
* install_formats() - May have to do this as well.

```{r, echo=TRUE}
library("rio")
```

### Using Rio to Import
There are other libraries to do this, but here's one suggestion from another faculty member.
```{r pressure, echo=TRUE}
URL <- "https://www.dropbox.com/s/130c5ol3o2jjmgk/public.fin.msm.model.csv?raw=1"
FinRegulatorData <- rio::import(URL)
```

### Structure
Let's take a look at the structure of the dataset.
```{r, echo=TRUE, include=TRUE}
summary(FinRegulatorData)
str(FinRegulatorData)
unique(FinRegulatorData$country)

# Change to factor
FinRegulatorData$country<-as.factor(FinRegulatorData$country)
```

### Generating counts of a category
The plyr library used to be one of the biggest libraries in R.  Here, we'll see how easy it is to count factors.
```{r, echo=TRUE, include=TRUE}
library('plyr')
count(FinRegulatorData, "country")
```

### Using the updated dplyr package
This looks extremely similar, but it offers information in tibble format.
```{r, echo=TRUE, include=TRUE}
library('dplyr')
count(FinRegulatorData, country)
```

