---
title: "Tidy Data - A Primer"
author: "Tom Tibbett"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Disclaimer: These are not real data.  They have been fabricated for instructional purposes.

### Live Coding Session Demo
First, let's read in, well, read_in.txt.  This file is tab delimited with no headers.  My client informed me what the column names are.

```{r, echo=TRUE}
df <- read.table("../../data/read_in.txt", sep="\t", header=FALSE, col.names = c("Name", "ID", "IQ", "Grades", "Legs"))
head(df)
```

Next, let's take a look at some of the summary information.

```{r, echo=TRUE}
summary(df)
unique(df$Legs)
```

Hmm... that's weird that someone has three legs.  I ought to go back and ask the client about what the Legs column means!

### Splitting a Name

We'll be using the reshape2 package as an alternative to the Lecture Assignments.  With R, you have options in what library you use.

I'm going to use colsplit(), because it's a quick way of cutting out names if there's an obvious separator.  Here, we have an underscore.

```{r, echo=TRUE}
library("reshape2")
vars<-colsplit(df$Name, "_", names=c("FirstName", "LastName"))
df<-cbind(vars, df) # Adding it back to the original dataframe.
head(df)
```

### Searching with grep

First off, remember that subsetting dataframes is as simple as ```df[{row}, {column}]```.

* df[1,] - if you want the first row and all columns
* df[1, c(1,2)] - if you want the first row and columns 1 and 2
* df[1, c("ID", "Gophers")] - Same deal, but with explicit names

You can even use logical statements to affect your dataframe.

* df[which(df$Legs==3),] - To find the row(s) which there are persons with three legs.
* df[which(df$Legs < 3),] - To find the row(s) which there are persons with less than three legs.
* df[order(df$IQ), ] - To sort your dataframe by ascending IQ score

To find a particular pattern within strings, you can format things similarly to the df[row, columns] structure. Using regular expression syntax, which I highly recommend you read up using ```?grep```, you can search for and display matches to patterns in your character data.

My client would like to see all the people with last names that start with "Pa."

* ```^``` indicates you want to make sure this is at the start of a string.
* ```[Pp]``` indicates the first character should be either a P or a p.
* ```a``` indicates the next letter should be a lowercase a.
* So, a match here indicates that a last name starts with either 'Pa' or 'pa.'

```{r, echo=TRUE}
df[grep('^[Pp]a', df$LastName),]
```

### Reports

Now I can put all of this together to create reports my client needs.

* Ordering by descending IQ
* Including First Name and Last Name
* Dropping all other variables

```{r}
report<-df[order(-df$IQ),c("FirstName", "LastName", "IQ")]
report
```