---
title: "Preparing Data for Analysis - A Primer"
author: "Tom Tibbett"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this demo, I'll be grabbing some data from online and turning it into something usable.  Make sure you have the relevant libraries installed and ready to go.

### Library Import

```{r, echo=TRUE}
library("rvest")   # Great for grabbing and parsing HTML
library("dplyr")   # Easy transformation of data.frames for summarization
library("tidyr")   # Nice way to arrange data
library("ggplot2") # Excellent for visuals
```

We are going to pull some data from the ESPN website.  Feel free to take a look at the URL below.

```{r, echo=TRUE}
url <- 'http://espn.go.com/nfl/superbowl/history/winners'
webpage<- read_html(url)
```

Using rvest, I'm going to grab some of the HTML in order to investigate nodes of interest: namely, the table.

```{r, echo=TRUE}
sb_table<-html_nodes(webpage, 'table')
sb<-html_table(sb_table)
```

Then, I'm going to turn this into a data frame we can (sort of) use.

```{r, echo=TRUE}
a<-data.frame(sb)
names(a)<-a[2,]
df<-slice(a, 3:length(a$RESULT))
df
```

### Regular Expressions

It looks like there are four potential variables in one column.  Team1, Team1's score, Team2, and Team2's score.  I can easily split the two of them with a comma using the ```separate()``` function in dplyr.  But the problem after that is how to separate Teams from their score. 

If you notice, there are some issues with simply splitting columns by spaces, like we did we Team1's variables and Team2's.  Teams like "Green Bay" have two spaces before their score.  We'll have to devise another strategy.  One way is regular expressions.  I highly recommend researching ```regex``` online.

Here, I do the following in order.

* [ ] - match a single character present in the set: here I chose a space.
* (?= ) - positive lookahead
* [^ ] - match a single character NOT present in the set: here I chose not a space
* \+ - matches at least once
* $ - asserts position at the end of the line
 
So, it says:

* Look for a space.
* Got it?  Now, right after that, look ahead for a not-space.
* Is it the last one on the line? If so, split it!

It will go character by character until it finds the last space of the string.

```{r, echo=TRUE}
clean<-df %>%
  separate(RESULT, into=c("Team1", "Team2"), sep=", ") %>%
  # After splitting this way, I want the last 'word' from each column
  separate(Team1, into=c("Team1", "Team1_Score"),sep="[ ](?=[^ ]+$)") %>%
  separate(Team2, into=c("Team2", "Team2_Score"),sep="[ ](?=[^ ]+$)")

clean
```

Now, let's make this a dataframe and alter the column names.

```{r, echo=TRUE}
df<-data.frame(clean)
names(df) <- c("SuperBowl", "Date", "Site", "Team1", "Team1_Score", "Team2", "Team2_Score")
head(df)
```

### Troubleshooting

One problem we have is that the data is not in the right format to be utilized for anything.  We need to change that.
```{r, echo=TRUE}
data.frame(sapply(df, class))
```

Fortunately, we can very easily coerce these variables to other types, given the format.  For our date variable, especially, it's important to look up the keys for month, day, and year formats and combine them.  Checking again, we see that our dataframe has changed to the appropriate type.

```{r, echo=TRUE}
df[,c("Team1_Score", "Team2_Score")]<-sapply(df[,c("Team1_Score", "Team2_Score")], as.numeric) # Our numeric variables.
df$Date<-as.Date(df$Date, format='%b. %d, %Y') # Our date variable.

data.frame(sapply(df, class))
```

Now that we've converted our columns to numeric, we can use ```dplyr``` and ```tidyr``` to process our data.

```{r, echo=TRUE}
df %>%
  count(Team1) %>%
  arrange(desc(n))
```

Always be ready for curveballs.  If we look at the ```unique()``` values, we can see that there are some discrepancies in recording.  Occasionally, there are only locations.  But sometimes, a team name is attached.  This is worth fact-checking later, but for the moment, we'll assume the records are correct and do not need editing.

```{r, echo=TRUE}
unique(df$Team1)
```

### Making Something from Nothing

My friend is unfamiliar with American football, and the Super Bowl is coming up.  He would like to watch five Super Bowls to amp himself up for his first live event.  However, he's not sure where to begin and asks me for help.  He wants to watch an exciting, competitive match.  He asks me to choose some good ones to watch.

I can use this data that I have created to visualize how competitive each Super Bowl is and create a spoiler-free list for him.  I decide to use quartiles to determine when a game is either a nail-biter or a blowout.  From the ```summary```, I can grab and create metrics from the Victory Spread: the distance in final score between Team1 and Team2.

```{r, echo=TRUE}
df["VictorySpread"]<- abs(df["Team1_Score"]-df["Team2_Score"]) # I use absolute value here because Team2 might outscore Team1.

summary(df$VictorySpread)
FQ<-as.numeric(summary(df$VictorySpread)[2]) # This is the first quartile that I grab
TQ<-as.numeric(summary(df$VictorySpread)[5]) # This is the third.
```

Using these metrics, I can create a column categorizing whether something is a close game, a standard one, or a blowout.

```{r, echo=TRUE}
df["WasItClose"]<-"Standard" # Make everything standard at first, then overwrite!
df$WasItClose[which(df$VictorySpread<=FQ)] <- "Nail Biter" # Lower than first quartile might be extremely competitive
df$WasItClose[which(df$VictorySpread>=TQ)] <- "Blowout" # Higher than the third quartile might be hilariously uncompetitive
```

Let's plot this.

* Because I converted the ```Date``` column to datetime format, it can be plotted on an axis!
* I use VictorySpread as the height of the bars.
* I color the bars by ```WasItClose```.
* I give an informative title.
* I rename the axes.

```{r, echo=TRUE, include=TRUE}
ggplot(df, aes(Date, VictorySpread))+
  geom_bar(aes(fill=WasItClose), stat="identity")+
  ggtitle("Super Bowls - Degrees of Competitive") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Super Bowl Year") +
  ylab("Final Score Discrepancy")
```

Finally, I can deliver the list using ```dplyr``` and ```tidyr``` piping.

* I ```select``` four of the columns I need.
* I ```filter``` the observations so that only Nail Biter games remain.
* I ```arrange``` the dataset by ascending Victory Spread; if there are ties, secondary order is by most recent.
* I ```slice``` the result so that only the Top Five remain.
* To remove spoilers, I un```select``` the variables I do not need anymore.

```{r, echo=TRUE}
watchthese<-df %>%
  select(SuperBowl,VictorySpread,WasItClose, Date) %>% # Select These Variables Only
  filter(WasItClose=="Nail Biter") %>% # Only consider Nail Biters
  arrange(VictorySpread,desc(Date)) %>% # Arrange them by score and how recent they were
  slice(1:5) %>% # Pick the closest - in event of a tie, prioritize most recent
  select(-WasItClose, -VictorySpread) # Remove the Spoilers

data.frame(watchthese)
```

